\LARGE \bf The Battle for Filter Supremacy: A Retrospective: The Movie: The Game: The Paper

### 2014-12-23 Report/project_report.tex
Another term was missing from F
Some minor changes relating to noise

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EXTENDED KALMAN FILTER} \label{sec:ekf}
In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of features by recursively updating the state as new measurements become available.
%Features are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
%In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SLIDING WINDOW FILTER} \label{sec:slidingwindow}
In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.

@@

\subsection{IMU State Estimate Propagation}
The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by a continuous-time motion model:
\begin{align}
    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} \\
    \dot{\hat{\Vec{b}}}_{\boldsymbol{\omega}}  &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T\left(\Vec{v}_m - \hat{\Vec{b}}_{\Vec{v}} \right)
\end{align}
where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$ and
\begin{align*}
    \boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right) &= \bbm -\hat{\boldsymbol{\omega}}^\times & \hat{\boldsymbol{\omega}} \\
                                                                -\hat{\boldsymbol{\omega}}^T & 0
                                                            \ebm 
\end{align*}
with
\begin{align*}
    \hat{\boldsymbol{\omega}} = \boldsymbol{\omega}_m - \hat{\Vec{b}}_{\boldsymbol{\omega}}
\end{align*}
and
\begin{align*}
    \hat{\boldsymbol{\omega}}^\times &=   \bbm    0 & -\hat{\omega}_3 & \hat{\omega}_2 \\
                                                \hat{\omega}_3 & 0 & -\hat{\omega}_1 \\
                                                -\hat{\omega}_2 & \hat{\omega}_1 & 0
                                            \ebm .
\end{align*}

In our implementation we propagate the motion model using a simple forward-Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.

@@

We obtain the predicted IMU-IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ by integrating
\begin{align} \label{eq:imucovar}
    \dot{\hat{\Vec{P}}}_{II}\left(t_k, t_k  + \tau \right) &= \Vec{F} \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \notag \\ 
                & + \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \Vec{F}^T \notag \\ 
                & + \Vec{G} \Vec{Q}_I \Vec{G}^T, \tau\in\left[0,T\right]
\end{align}
with the initial condition $\hat{\Vec{P}}_{II}\left(t_k, t_k\right) = \hat{\Vec{P}}_{II,k}$. 
The Jacobians $\Vec{F}$, $\Vec{G}$ are given by
\begin{align}
\Vec{F} &= \bbm -\hat{\boldsymbol{\omega}}^\times & -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3} \\
            \ebm
\end{align}
\begin{align}
\Vec{G} &=  \bbm    -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                                \Vec{0}_{3\times3} & \Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{1}_3 \\
                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3}
            \ebm ,
\end{align}
and $\Vec{Q}_I$ is the covariance of the IMU measurement noise.

@@

If we linearize (\ref{eq:residual_def}) about the estimates for the camera pose and feature position, we obtain an estimate of the exteroceptive measurement error
\begin{align}
    \Vec{r}_i^{(j)} &\simeq \Vec{H}_{\Vec{x}, i}^{(j)} \widetilde{\Vec{x}}_i + \Vec{H}_{f, i}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}_i^{(j)}
\end{align}
where $\Vec{H}_{\Vec{x}, i}^{(j)}$ and $\Vec{H}_{f, i}^{(j)}$ are the Jacobians of the measurement of feature $f_j$ from camera pose $i$ with respect to the filter state and the position of the feature, respectively.
These are given by
\begin{align}
    \Vec{H}_{\Vec{x}, i}^{(j)} &= \bbm \Vec{0} & \Vec{J}^{\left(j\right)}_i \left(\hat{\Vec{p}}_{C_i}^{f_j C_i}\right)^\times & -\Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G} & \Vec{0} \ebm \\
    \Vec{H}_{f, i}^{(j)} &= \Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G}
\end{align}
where the left $\Vec{0}$ in $\Vec{H}_{\Vec{x}, i}^{(j)}$ has dimension $2\times\left(12+6\left(i-1\right)\right)$, the right $\Vec{0}$ has dimension $2\times6\left(N-i\right)$, and
\begin{align}
    \Vec{J}^{\left(j\right)}_i &= \frac{1}{\left(\hat{Z}^{(j)}_{C_i}\right)^2}
    \bbm \hat{Z}^{(j)}_{C_i} & 0 & -\hat{X}^{(j)}_{C_i} \vspace{0.5em} \\
             0 & \hat{Z}^{(j)}_{C_i} & -\hat{Y}^{(j)}_{C_i}
    \ebm .
\end{align}
$\Vec{n}_i^{(j)}$ is a zero-mean Gaussian noise term with covariance matrix $\Vec{R}_i^{(j)} = \sigma^2_{\text{im}} \Vec{1}_{2}$.

By stacking the errors $\Vec{r}_i^{(j)}$, we arrive at an expression for the $2M_j \times 1$ measurement error vector of feature $f_j$ over the entire window of camera poses, where $M_j$ is the number of camera poses from which feature $f_j$ was observed:
\begin{align} \label{eq:residual_stacked}
    \Vec{r}^{(j)} &\simeq \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}^{(j)}
\end{align}
However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the error in the state.
Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly. [TALK HERE ABOUT WHAT HAPPENS IF YOU LEAVE IT IN]

@@

Since $\Vec{H}_{f}^{(j)}$ has full column rank, $\Vec{A}$ has dimension $2M_j \times \left(2M_j - 3\right)$ and $\Vec{r}_o^{(j)}$ has dimension $\left(2M_j - 3\right) \times 1$.
Furthermore, because $\Vec{A}$ is semi-unitary, the covariance matrix of $\Vec{n}_o^{(j)}$ is given by $\Vec{R}_o = \sigma^2_{\text{im}}\Vec{A}^T\Vec{A} = \sigma^2_{\text{im}}\Vec{1}_{2M_j - 3}$.

We can now stack all the errors $\Vec{r}_o^{(j)}$ for all the features in the current batch to arrive at
\begin{align} \label{eq:residual_nullspace_stacked}
    \Vec{r}_o &= \Vec{H}_o \widetilde{\Vec{x}} + \Vec{n}_o .
\end{align}
The dimension of this vector can be quite large in practice, so we use the QR-decomposition of $\Vec{H}_o$ to reduce the computational complexity of the EKF update:
\begin{align}
    \Vec{H}_o = \bbm \Vec{Q}_1 & \Vec{Q}_2 \ebm \bbm \Vec{T}_\Vec{H} \\ \Vec{0} \ebm
\end{align}
where $\Vec{Q}_1, \Vec{Q}_2$ are unitary matrices and $\Vec{T}_\Vec{H}$ is an upper-triangular matrix.
Substituting this result into (\ref{eq:residual_nullspace_stacked}) and pre-multiplying by $\bbm \Vec{Q}_1 & \Vec{Q}_2 \ebm ^T$, we obtain
\begin{align}
    \bbm \Vec{Q}_1^T \Vec{r}_o \\ \Vec{Q}_2^T \Vec{r}_o \ebm &= \bbm \Vec{T}_\Vec{H} \\ \Vec{0} \ebm \widetilde{\Vec{x}} + \bbm \Vec{Q}_1^T \Vec{n}_o \\ \Vec{Q}_2^T \Vec{n}_o \ebm .
\end{align}
Noting that the quantity $\Vec{Q}_2^T \Vec{r}_o$ is only noise, we discard it and define a new error term that we use in the EKF update:
\begin{align}
    \Vec{r}_n &:= \Vec{Q}_1^T \Vec{r}_o = \Vec{T}_\Vec{H} \widetilde{\Vec{x}} + \Vec{Q}_1^T \Vec{n}_o =: \Vec{T}_\Vec{H} \widetilde{\Vec{x}} + \Vec{n}_n
\end{align} 
The covariance matrix of $\Vec{n}_n$ is given by $\Vec{R}_n = \Vec{Q}_1^T \Vec{R}_o \Vec{Q}_1 = \sigma^2_\text{im} \Vec{1}_r$ where $r$ is the number of columns in $\Vec{Q}_1$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Extended Kalman Filter (EKF)} \label{subsec:ekf}
In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of features by recursively updating the state as new measurements become available.
%Features are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
%In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sliding Window Filter (SWF)} \label{subsec:slidingwindow}
In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.

@@

\subsection{IMU State Estimate Propagation}
The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by a continuous-time motion model:
\begin{align}
    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} \\
    \dot{\hat{\Vec{b}}}_{\boldsymbol{\omega}}  &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}
\end{align}
where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$,
\begin{align*}
    \hat{\Vec{v}} &= \Vec{v}_m - \hat{\Vec{b}}_{\Vec{v}}, & \hat{\boldsymbol{\omega}} &= \boldsymbol{\omega}_m - \hat{\Vec{b}}_{\boldsymbol{\omega}},
\end{align*}
\begin{align*}
    \boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right) &= \bbm -\hat{\boldsymbol{\omega}}^\times & \hat{\boldsymbol{\omega}} \\
                                                                -\hat{\boldsymbol{\omega}}^T & 0
                                                            \ebm ,
                                                            & \text{and} &&
    \hat{\boldsymbol{\omega}}^\times &=   \bbm    0 & -\hat{\omega}_3 & \hat{\omega}_2 \\
                                                \hat{\omega}_3 & 0 & -\hat{\omega}_1 \\
                                                -\hat{\omega}_2 & \hat{\omega}_1 & 0
                                            \ebm .
\end{align*}
In our implementation we propagate the motion model using a simple forward-Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.

We can also examine the linearized continuous-time model of the IMU error state:
\begin{align}
    \dot{\widetilde{\Vec{x}}}_I = \Vec{F}\widetilde{\Vec{x}}_I + \Vec{G}\Vec{n}_I
\end{align}
where the Jacobians $\Vec{F}$, $\Vec{G}$ are given by
\begin{align}
\Vec{F} &= \bbm -\hat{\boldsymbol{\omega}}^\times & -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                    -\hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}^\times & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3} \\
            \ebm
\end{align}
\begin{align}
\Vec{G} &=  \bbm    -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                                \Vec{0}_{3\times3} & \Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{1}_3 \\
                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3}
            \ebm ,
\end{align}
and $\Vec{n}_I = \bbm \Vec{n}_{\boldsymbol{\omega}}^T & \Vec{n}_{\Vec{b}_{\boldsymbol{\omega}}}^T & \Vec{n}_{\Vec{v}}^T & \Vec{n}_{\Vec{b}_{\Vec{v}}}^T \ebm^T$ is the IMU process noise, which has covariance matrix $\Vec{Q}_I$.

@@

We obtain the predicted IMU-IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ by integrating
\begin{align} \label{eq:imucovar}
    \dot{\hat{\Vec{P}}}_{II}\left(t_k, t_k  + \tau \right) &= \Vec{F} \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \notag \\ 
                & + \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \Vec{F}^T \notag \\ 
                & + \Vec{G} \Vec{Q}_I \Vec{G}^T, \tau\in\left[0,T\right]
\end{align}
with the initial condition $\hat{\Vec{P}}_{II}\left(t_k, t_k\right) = \hat{\Vec{P}}_{II,k}$.

@@

If we linearize (\ref{eq:residual_def}) about the estimates for the camera pose and feature position, we obtain an estimate of the exteroceptive measurement error
\begin{align}
    \Vec{r}_i^{(j)} &\simeq \Vec{H}_{\Vec{x}, i}^{(j)} \widetilde{\Vec{x}}_i + \Vec{H}_{f, i}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}_i^{(j)}
\end{align}
where $\Vec{H}_{\Vec{x}, i}^{(j)}$ and $\Vec{H}_{f, i}^{(j)}$ are the Jacobians of the measurement of feature $f_j$ from camera pose $i$ with respect to the filter state and the position of the feature, respectively.
These are given by
\begin{align}
    \Vec{H}_{\Vec{x}, i}^{(j)} &= \bbm \Vec{0} & \Vec{J}^{\left(j\right)}_i \left(\hat{\Vec{p}}_{C_i}^{f_j C_i}\right)^\times & -\Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G} & \Vec{0} \ebm \\
    \Vec{H}_{f, i}^{(j)} &= \Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G}
\end{align}
where the left $\Vec{0}$ in $\Vec{H}_{\Vec{x}, i}^{(j)}$ has dimension $2\times\left(12+6\left(i-1\right)\right)$, the right $\Vec{0}$ has dimension $2\times6\left(N-i\right)$, and
\begin{align}
    \Vec{J}^{\left(j\right)}_i &= \frac{1}{\left(\hat{Z}^{(j)}_{C_i}\right)^2}
    \bbm \hat{Z}^{(j)}_{C_i} & 0 & -\hat{X}^{(j)}_{C_i} \vspace{0.5em} \\
             0 & \hat{Z}^{(j)}_{C_i} & -\hat{Y}^{(j)}_{C_i}
    \ebm .
\end{align}
$\Vec{n}_i^{(j)}$ is a zero-mean Gaussian noise term with covariance matrix $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_u, \sigma^2_v\right\}$.

By stacking the errors $\Vec{r}_i^{(j)}$, we arrive at an expression for the $2M_j \times 1$ measurement error vector of feature $f_j$ over the entire window of camera poses, where $M_j$ is the number of camera poses from which feature $f_j$ was observed:
\begin{align} \label{eq:residual_stacked}
    \Vec{r}^{(j)} &\simeq \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}^{(j)}
\end{align}
The noise vector $\Vec{n}^{(j)}$ has covariance matrix $\Vec{R}^{(j)} = \text{diag}\left\{\Vec{R}_1^{(j)}, \hdots, \Vec{R}_{M_j}^{(j)}\right\}$.

However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the error in the state.
Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly. [TALK HERE ABOUT WHAT HAPPENS IF YOU LEAVE IT IN]

@@

Since $\Vec{H}_{f}^{(j)}$ has full column rank, $\Vec{A}$ has dimension $2M_j \times \left(2M_j - 3\right)$ and $\Vec{r}_o^{(j)}$ has dimension $\left(2M_j - 3\right) \times 1$.
The covariance matrix of $\Vec{n}_o^{(j)}$ is given by $\Vec{R}_o^{(j)} = \Vec{A}^T\Vec{R}^{(j)}\Vec{A}$.

We can now stack all the errors $\Vec{r}_o^{(j)}$ for all the features in the current batch to arrive at
\begin{align} \label{eq:residual_nullspace_stacked}
    \Vec{r}_o &= \Vec{H}_o \widetilde{\Vec{x}} + \Vec{n}_o .
\end{align}
The dimension of this vector can be quite large in practice, so we use the QR-decomposition of $\Vec{H}_o$ to reduce the computational complexity of the EKF update:
\begin{align}
    \Vec{H}_o = \bbm \Vec{Q}_1 & \Vec{Q}_2 \ebm \bbm \Vec{T}_\Vec{H} \\ \Vec{0} \ebm
\end{align}
where $\Vec{Q}_1, \Vec{Q}_2$ are unitary matrices and $\Vec{T}_\Vec{H}$ is an upper-triangular matrix.
Substituting this result into (\ref{eq:residual_nullspace_stacked}) and pre-multiplying by $\bbm \Vec{Q}_1 & \Vec{Q}_2 \ebm ^T$, we obtain
\begin{align}
    \bbm \Vec{Q}_1^T \Vec{r}_o \\ \Vec{Q}_2^T \Vec{r}_o \ebm &= \bbm \Vec{T}_\Vec{H} \\ \Vec{0} \ebm \widetilde{\Vec{x}} + \bbm \Vec{Q}_1^T \Vec{n}_o \\ \Vec{Q}_2^T \Vec{n}_o \ebm .
\end{align}
Noting that the quantity $\Vec{Q}_2^T \Vec{r}_o$ is only noise, we discard it and define a new error term that we use in the EKF update:
\begin{align}
    \Vec{r}_n &:= \Vec{Q}_1^T \Vec{r}_o = \Vec{T}_\Vec{H} \widetilde{\Vec{x}} + \Vec{Q}_1^T \Vec{n}_o =: \Vec{T}_\Vec{H} \widetilde{\Vec{x}} + \Vec{n}_n
\end{align} 
The covariance matrix of $\Vec{n}_n$ is given by $\Vec{R}_n = \Vec{Q}_1^T \Vec{R}_o^{(j)} \Vec{Q}_1$.
